---
title: "Problem Set 1: Bayesian Inference"
subtitle: By Becks Lopez
output:
  pdf_document: default
  html_document:
    df_print: paged
---
#Setting up Necessary Packages 
```{r}
library(rethinking)
```

# Problem 2M1: Globe Tossing
We assume a uniform distrubtion prior for $p$ which is the true proportion of water coving the globe.And consider three different cases for posterior distribution:

(1)W,W,W
```{r}
#Set up grid 
p_grid<-seq(0,1,length.out = 100) 

#Likelihood of getting 3 W within 3 tosses
likelihood<-dbinom(3,size=3,prob=p_grid)

prior <-  rep(1,100)
posterior <- likelihood*prior
posterior <- posterior/sum(posterior)

#Produce plot of posterior distribution
plot(posterior ~ p_grid,type="l", main="(1) W W W")

```
(2) W,W,W,L
```{r}
#Set up grid 
p_grid<-seq(0,1,length.out = 100) 

#Likelihood of getting 3 W within 4 tosses
likelihood2<-dbinom(3,size=4,prob=p_grid)

prior <-rep(1,100)
posterior2 <- likelihood2*prior
posterior2 <- posterior2/sum(posterior2)
#Produce plot of posterior distribution
plot(posterior2 ~ p_grid,type="l", main="(2) W W W L")
```
(3)L,W,W,L,W,W,W
```{r}
#Set up grid 
p_grid<-seq(0,1,length.out = 100) 

#Likelihood of getting 5 W within 7 tosses
likelihood3<-dbinom(5,size=7,prob=p_grid)

prior <-  rep(1,100)
posterior3 <- likelihood3*prior
posterior3 <- posterior3/sum(posterior3)
#Produce plot of posterior distribution
plot(posterior3 ~ p_grid,type="l", main="(3)L,W,W,L,W,W,W")
```

Notice that with increased data size, we begin to convergence towards the true $p$.

# Problem 2M2:
Here we assume that our prior equals zero for $p<0.5$ and is a positive constant for $p \geq 0.5$ and repeat the process from problem 2M1 for three cases.

(1)W,W,W
```{r}
p_grid <-seq(0,1,length.out = 100)
prior_truncated <-ifelse(p_grid<0.5,0,1)
posterior_truncated <-likelihood*prior_truncated
posterior_truncated <-posterior_truncated/sum(posterior_truncated)
plot(posterior_truncated ~ p_grid,type="l", main="(1) W W W")
```


(2) W,W,W,L
```{r}
posterior_truncated2 <-likelihood2*prior_truncated
posterior_truncated2 <-posterior_truncated2/sum(posterior_truncated2)
plot(posterior_truncated2 ~ p_grid,type="l", main="(2) W W W L")
```
(3)L,W,W,L,W,W,W
```{r}
posterior_truncated3 <- likelihood3*prior_truncated
posterior_truncated3 <- posterior_truncated3/sum(posterior_truncated3)
plot(posterior_truncated3 ~ p_grid,type="l", main="(3)L,W,W,L,W,W,W")
```
# Problem 2M3
We know that earth is 70% water which implies 30% land. Additionally, we are given that Mars is 100% land, and 0% water.Therefore, we know that: 
$$P(\text{land}|\text{Earth})=0.3\\
P(\text{land}|\text{Mars})=1.0$$
We assume that each globe was equally likely to be tossed which implies that $P(\text{Mars})=P(\text{Earth})=0.5.$ We know that based on the rules of conditional probability that
$$P(\text{Earth}|\text{Land})=\frac{P(\text{land}|\text{Earth})P(\text{Earth})}{P(\text{Land})}$$ where $P(\text{Land})=P(\text{land}|\text{Earth})*P(\text{Earth})+P(\text{land}|\text{Mars})*P(\text{Mars}).$ We can substitute in our probabilitys to state:
$$P(\text{Earth}|\text{Land})=\frac{P(\text{land}|\text{Earth})P(\text{Earth})}{P(\text{land}|\text{Earth})*P(\text{Earth})+P(\text{land}|\text{Mars})*P(\text{Mars})}\\
= \frac{0.3*0.5}{0.3*0.5+1*0.5}.$$ We compute below to find that $P(\text{Earth}|\text{Land})=0.23.$

```{r}
P_EL=0.3*0.5 / ( 0.3*0.5 + 1*0.5 )
P_EL
```

#Chapter3 Problems
Blow is the code from the textbook to set up our posterior distrubtion for problems 3E1-3E7.
```{r}
#Code for #3E1-3E7 from textbook
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```

```{r}
#Probem 3E1: How much posterior probability lies below p = 0.2?
sum(samples<0.2)/1e4

#Problem 3E2: How much posterior probability lies above p = 0.8?
sum(samples>0.8)/1e4

#Problem 3E3: How much posterior probability lies between p = 0.2 and p = 0.8?
sum(samples>0.2 & samples <0.8)/1e4

#Problem3E4: 20% of the posterior probability lies below which value of p?
quantile(samples, 0.2)

#Problem3E5: 20% of the posterior probability lies above which value of p?
quantile(samples, 0.8)

#Problem3E6: Which values of p contain the narrowest interval equal to 66% of the posterior probability?
HPDI(samples, prob =0.66)

#Problem3E7: Which values of p contain 66% of the posterior probability, assuming equal posterior probability both below and above the interval?
PI(samples, prob=0.66)

```
#Chapter 3 Problems:

#Problem 3M1:
We return to the globe tossing example and construct our posterior distribution using a grid approximation assuming a flat prior.

```{r}
#Code utilized from textbook
p_grid <- seq(0,1 ,length.out=1000 )
prior <- rep(1,1000)

likelihood <- dbinom(8,size=15 , prob=p_grid )
posterior <- likelihood * prior_truncated

posterior <- posterior / sum(posterior)
plot(posterior ~ p_grid,type="l", main="8 W in 15 Tosses")
```

#Problem 3M2:
We draw 10000 samples from our grid approximnation and the sames to find the narrowest interval equal to $90 \%$ of our posterior probability.
```{r}
samples <- sample( p_grid , size=1e4 , replace=TRUE , prob=posterior )
HPDI(samples, prob =0.90)
```

#Problem 3M3: 
We construct a posterior predictive check for our model where we generate random binomial samples and find the probability of observin 8 waters in 15 tosess:
```{r}
w<-rbinom(1e4,size = 15, prob=samples)
sum(w==8)/1e4
```

#Problem 3M4
We utiilize the same samples in order to utilize the posterior distribution as before in order to find the probability of observing 6 water in 9 tosses:
```{r}
w<-rbinom(1e4,size = 9, prob=samples)
sum(w==6)/1e4
```

#Problem 3M5
We rerun through our code above but with one difference, we now use a prior that is zero below our threshold $(p=0.5)$ and a constant above that threshold:
```{r}
#Construct the posterior distribution, using grid approximation. 
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior_truncated <-ifelse(p_grid<0.5,0,1)
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior_truncated <- likelihood * prior_truncated
posterior_truncated <- posterior_truncated / sum(posterior_truncated)
plot(posterior_truncated ~ p_grid,type="l", main="8 W in 15 Tosses")

# Draw 10,000 samples from the grid approximation from above. Then use the samples to calculate the 90% HPDI for p
samples <- sample( p_grid , size=1e4 , replace=TRUE , prob=posterior )
HPDI(samples, prob =0.90)

#Construct a posterior predictive check for this model and data. Find the probability of observing 8 water in 15 tosses.
w<-rbinom(1e4,size = 15, prob=samples)
sum(w==8)/1e4

#Using the posterior distribution constructed, find the probability of observing 6 water in 9 tosses.
w<-rbinom(1e4,size = 9, prob=samples)
sum(w==6)/1e4
```

Note that using the prior with a threshold actually ended up giving us higher probabilities for 6 W in 15 tosses (0.1851) versus a flat prior (0.1735) where $6/15=0.\bar{6}$ which is close to our true value of $p=0.7.$ We see that while $8/15=0.5\bar{3}$ is slighly further from our true value and thus has lower probability, the prior with a threhold still makes a better inference (0.1486) than the flat uniform prior (0.1479).

